---
title: "Reproducible Research: Project 1"
author: "Ryan Honomichl"
date: "August 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries; loading and preprocessing the data

```{r, echo=TRUE, message = FALSE}
library(dplyr)
library(ggplot2)
library(mice)
```

```{r, echo=TRUE, message = FALSE}

if(dir.exists("C:/Users/Ryan Honomichl/Dropbox/Coursera/5_Reproducible_Research/Week_2/project")) { 
  
  setwd("C:/Users/Ryan Honomichl/Dropbox/Coursera/5_Reproducible_Research/Week_2/project")
  
}else{
  
  dir.create("C:/Users/Ryan Honomichl/Dropbox/Coursera/5_Reproducible_Research/Week_2/project")
  
  setwd("C:/Users/Ryan Honomichl/Dropbox/Coursera/5_Reproducible_Research/Week_2/project")
}

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(fileUrl,destfile="./Dataset.zip")

unzip(zipfile="Dataset.zip")

activity.raw <- read.csv("./activity.csv", header=TRUE, sep=",")

```

### Check structure of data frame

```{r, echo=TRUE}
str(activity.raw)
```

NOTE: date as factor

```{r, echo=TRUE}
activity.raw$date <- as.Date(activity.raw$date, "%Y-%m-%d")
```

```{r, echo=TRUE}
str(activity.raw)
```

Note: Now date is a date variable

## Calculate the total number of steps taken per day (ignoring NAs)

```{r, echo=TRUE}
steps.sum <- activity.raw %>% 
  filter(!is.na(steps)) %>%
  group_by(date) %>% 
  summarize(TotalSteps=sum(steps))

print(steps.sum)
```

## Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
s <- ggplot(steps.sum, aes(x = TotalSteps))  

s + geom_histogram(fill = "blue",binwidth = 1000) +   labs(title = "Total Steps (by day)", x = "Total Steps",
  y = "Frequency")
```

##Calculate and report the mean and median of the total number of steps taken per day

```{r, echo=TRUE}
mean(steps.sum$TotalSteps)

median(steps.sum$TotalSteps)
```

##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
act.no.na <- na.omit(activity.raw)

fivemin.int <- act.no.na %>%
  group_by(interval) %>%
  summarize(aver.steps=mean(steps))

s2 <- ggplot(fivemin.int, aes(x = interval, y = aver.steps)) 

s2 + geom_line() + labs(title = "Average Steps Taken (by 5 min. interval)", x = "Interval", y = "Average Steps")
```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE, Warnings = FALSE}
max_interv <- fivemin.int[which.max(fivemin.int$steps),1]

fivemin.int$interval[which.max(fivemin.int$aver.steps)]
```

##Imputing missing values: examine the number of missing data on the steps variable

```{r, echo=TRUE}
md.pattern(activity.raw)
```

NOTE: There are 15264 complete observations, 2304 observations with missing steps data

##Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo=TRUE}
imputed.data <- activity.raw

for (i in 1:nrow(imputed.data)) {
  if (is.na(imputed.data$steps[i])) {
    # Find the index value for when the interval matches the average
    temp.d <- which(imputed.data$interval[i] == fivemin.int$interval)
    # Assign the value to replace the NA
    imputed.data$steps[i] <- fivemin.int[temp.d,]$aver.steps
  }
}
```

####Calculate and report the mean and median total number of steps taken per day.  
####Do these values differ from the estimates from the first part of the assignment? What is the 
####impact of imputing missing data on the estimates of the total daily number of steps?

```{r, echo=TRUE}
steps.sum.imp <- imputed.data %>% 
  group_by(date) %>% 
  summarize(TotalSteps=sum(steps))

mean(steps.sum.imp$TotalSteps)

median(steps.sum.imp$TotalSteps)

```

NOTE: mean is unchanged, but median is slightly changed and is now equal to the mean.

##Make a histogram of the total number of steps taken each day 

```{r, echo=TRUE}
s3 <- ggplot(steps.sum.imp, aes(x = TotalSteps))  

s3 + geom_histogram(fill = "blue",binwidth = 1000) +   labs(title = "Total Steps (by day)", x = "Total Steps", y = "Frequency")

```

##Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether
##a given date is a weekday or weekend day.

```{r, echo=TRUE}

imputed.data$day <- weekdays(imputed.data$date)

imputed.data$daytype[imputed.data$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")] <- "weekday"

imputed.data$daytype[imputed.data$day %in% c("Saturday", "Sunday")] <- "weekend"

imputed.data$daytype <- as.factor(imputed.data$daytype)

```

####Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether
####a given date is a weekday or weekend day.

```{r, echo=TRUE}

imputed.data$day <- weekdays(imputed.data$date)

imputed.data$daytype[imputed.data$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")] <- "weekday"

imputed.data$daytype[imputed.data$day %in% c("Saturday", "Sunday")] <- "weekend"

imputed.data$daytype <- as.factor(imputed.data$daytype)

```

##Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
##and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r, echo=TRUE}
fivemin.int.imp <- imputed.data %>%
  group_by(daytype, interval)%>%
  summarize(aver.steps=mean(steps))

s4 <- ggplot(fivemin.int.imp, aes(x = interval, y = aver.steps)) 

s4 + geom_line() + facet_grid(daytype ~ .) + labs(title = "Average Steps Taken (by 5 min. interval and day type)",
    x = "Interval", y = "Average Steps")
```






























































































